Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-04-27 paper.tex

\subsection{Tree inference from clusters of samples}
[Discuss the "local" tree inference, but be careful not to use the term]

>>>>

\subsection{Tree inference from clusters of samples}
For a given daily batch of sample sequences that are to be attached to the
ARG, the LS HMM provides the most likely copying path and required mutations
for each of those sequences. We could directly attach these samples to
the ARG according to the copying paths and sets of mutations, but
such an approach would be highly unparsimonious because there are
often many samples that share the same copying path with
complex patterns of shared mutations.
In this section we discuss the strategy used to infer likely tree
toplogies among the samples that share the same copying path.
In the majority of cases this copying path will consist of a single
node (no recombination), and so for simplicity we will assume in the
following discussion that all of the samples in a HMM-defined cluster
attach to a single node. The differences required to deal with
newly-detected putative recombinants are discussed below
in Section XXX.

With tens of thousands of samples being added to the ARG per day,
such clusters of sequences attaching to the same node
can be large, with complex patterns of shared mutations.
To avoid the implied large polytomy and unparsimonious
repetition of identical mutations among sibling nodes, we
infer a tree among these samples (the LS HMM has already

A natural
way to

We can think of the sequences attaching to a given node in the ARG
from a daily batch as a polytomy,
but with potentially shared mutations among the sibling nodes
(Figure~\ref{fig:overview_sc2ts}C).
